Cosmological Hydrodynamic Simulations of Preferential Accretion in the SMBH of Milky Way Size Galaxies

### 2017-01-24 ms_predec282016.tex
Bolded for editor

\title{Preferential Accretion in the SMBH of Milky Way Size Galaxies \textcolor{green}{Due to Direct Feeding by Satellites}}
%\title{A Particular Appetite: Cosmological Hydrodynamic Simulations of Preferential Accretion in the Supermassive Black Holes of Milky Way Size Galaxies}

@@

Using a new, high-resolution cosmological hydrodynamic simulation of a Milky Way-type (MW-type) galaxy, we explore how a merger-rich assembly history affects the mass budget of the central supermassive black hole (SMBH). We examine a MW-mass halo at the present epoch whose evolution is characterized by several major mergers to isolate the importance of merger history on black hole accretion. This study is an extension of Bellovary et. al. 2013, which analyzed the accretion of high mass, high redshift galaxies and their central black holes, and found that the gas content of the central black hole reflects what is accreted by the host galaxy halo. In this study, we find that a merger-rich galaxy will have a central SMBH preferentially fed by merger gas. Moreover, we find that nearly 30$\%$ of the accreted mass budget of the SMBH enters the galaxy \textcolor{green}{directly from satellites during} the two major mergers in its history, which may account for the increase of merger-gas fueling the SMBH. Through an investigation of the angular momentum of the gas entering the host and its SMBH, we determine that merger gas enters the galaxy with lower angular momentum compared to smooth accretion, partially accounting for the preferential fueling witnessed in the SMBH. In addition, the presence of mergers, particularly major mergers, also helps funnel low angular momentum gas more readily to the center of the galaxy. Our results imply that galaxy mergers play an important role in feeding the SMBH in MW-type galaxies with merger-rich histories.

@@

Our simulation introduces a uniform UV background at z$\sim$9 to simulate the cosmic reionization energy energy using the formula of \cite{Haardt2012}. To model star formation, we stochastically turn gas particles into stars with a star formation efficiency parameter of c$^*$ = 0.1 once the density threshold and temperature satisfy conditions for star formation (10.0 amu cm$^{-3}$; T \textless 10$^4$ K).  \textcolor{green}{As shown in \citet{Governato10}, this high density threshold is necessary to produce bursty star formation events in the high-density peaks of the ISM.  Realistic star formation histories result in galaxies which have realistic density profiles and lie on the observed scaling relations (mentioned in detail below).  The lower-resolution simulations mentioned in this paper have a lower threshold, since the high density peaks are not resolved.}
These star particles represent a Kroupa initial mass function \citep{Kroupa1993}.  Molecular hydrogen and metal-line cooling are not included, though we implement a low-temperature extension to the cooling curve to trace metals \citep{Bromm2001} and the metal diffusion prescription of \cite{Shen2010}.

Supernova (SN) feedback releases $10^{51}$ ergs of thermal energy within a ``blastwave'' radius determined by the equations of \cite{Ostriker1988}. In the affected region, cooling turns off for a time relative to the expansion phase of the SN remnant also determined by the blastwave equation. SN Ia and II rates from \cite{Thielemann1986} and \cite{Woosley1995}, respectively, are implemented through the \cite{Raiteri1996} method, which uses the stellar lifetime calculations of the Padova group \citep{Alongi1993, Bressan1993, Bertelli1994} to describe stars with varying metallicities. Both the supernova ``blastwave'' radius and supernova (Ia and II) prescriptions are described in detail by \cite{Stinson2006}. \textcolor{violet}{While \textcolor{green}{it is true that} a different treatment of the \textcolor{green}{ISM and }SNe feedback might alter the structure of the ISM, \cite{Christensen2014} examined simulated spiral and dwarf galaxies utilizing similar SNe prescriptions \textcolor{green}{and three different ISM models,} and determined that the resulting ISM remained consistent with each other. Additionally, our galaxy is in good agreement with galaxies affected by superbubble SN feedback \citep{Keller2014}.}

>>>>

\title{Cosmological Hydrodynamic Simulations of Preferential Accretion in the SMBH of Milky Way Size Galaxies}
%\title{A Particular Appetite: Cosmological Hydrodynamic Simulations of Preferential Accretion in the Supermassive Black Holes of Milky Way Size Galaxies}

@@

Using a new, high-resolution cosmological hydrodynamic simulation of a Milky Way-type (MW-type) galaxy, we explore how a merger-rich assembly history affects the mass budget of the central supermassive black hole (SMBH). We examine a MW-mass halo at the present epoch whose evolution is characterized by several major mergers to isolate the importance of merger history on black hole accretion. This study is an extension of Bellovary et. al. 2013, which analyzed the accretion of high mass, high redshift galaxies and their central black holes, and found that the gas content of the central black hole reflects what is accreted by the host galaxy halo. In this study, we find that a merger-rich galaxy will have a central SMBH preferentially fed by merger gas. Moreover, we find that nearly 30$\%$ of the accreted mass budget of the SMBH enters the galaxy through the two major mergers in its history, which may account for the increase of merger-gas fueling the SMBH. Through an investigation of the angular momentum of the gas entering the host and its SMBH, we determine that merger gas enters the galaxy with lower angular momentum compared to smooth accretion, partially accounting for the preferential fueling witnessed in the SMBH. In addition, the presence of mergers, particularly major mergers, also helps funnel low angular momentum gas more readily to the center of the galaxy. Our results imply that galaxy mergers play an important role in feeding the SMBH in MW-type galaxies with merger-rich histories.

@@

Our simulation introduces a uniform UV background at z$\sim$9 to simulate the cosmic reionization energy energy using the formula of \cite{Haardt2012}. To model star formation, we stochastically turn gas particles into stars with a star formation efficiency parameter of c$^*$ = 0.1 once the density threshold and temperature satisfy conditions for star formation (10.0 amu cm$^{-3}$; T \textless 10$^4$ K). These star particles represent a Kroupa initial mass function \citep{Kroupa1993}. Molecular hydrogen and metal-line cooling are not included, though we implement a low-temperature extension to the cooling curve to trace metals \citep{Bromm2001} and the metal diffusion prescription of \cite{Shen2010}.

Supernova (SN) feedback releases $10^{51}$ ergs of thermal energy within a ``blastwave'' radius determined by the equations of \cite{Ostriker1988}. In the affected region, cooling turns off for a time relative to the expansion phase of the SN remnant also determined by the blastwave equation. SN Ia and II rates from \cite{Thielemann1986} and \cite{Woosley1995}, respectively, are implemented through the \cite{Raiteri1996} method, which uses the stellar lifetime calculations of the Padova group \citep{Alongi1993, Bressan1993, Bertelli1994} to describe stars with varying metallicities. Both the supernova ``blastwave'' radius and supernova (Ia and II) prescriptions are described in detail by \cite{Stinson2006}. \textcolor{violet}{While it's true a different treatment of the SNe feedback might alter the structure of the ISM, \cite{Christensen2014} examined simulated spiral and dwarf galaxies utilizing similar SNe prescriptions and determined that the resulting ISM remained consistent with each other. Additionally, our galaxy is in good agreement with galaxies affected by superbubble SN feedback \citep{Keller2014}.}
